A Switching Planner for Combined Task and Observation Planning

### 2010-11-27 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
STRIPS section dealt with

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. The underlying process
dynamics are modelled in terms of a finite set of stochastic actions
$\stochActions$, deterministic outcomes $\detActions$, and
state-characterising propositions $\propositions$.

We model a process state $\state$ as the set of propositions that are
true of the state. Notationally, we have $\state \subseteq
\props$. State change is induced by application of actions. A
stochastic action $\stochAction \in \stochActions$ is applicable if
its precondition $\poss(\stochAction)$, a set of propositions, are
satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
\state$. We write $\stochActions(\state)$ for set of actions
applicable at state $\state$.  If action $\stochAction \in
\stochActions(\state)$ is applied at \state, nature chooses one
element amongst a small set of deterministic outcomes
$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
nature takes outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
effect on the state given in terms of two lists of propositions. The
add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
$\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
..,\prop_m]$ is chosen by nature when \stochAction\ is
applied at state $\state$, then the resultant state is $ ( \state \cup
\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
propositions from $\add(\detAction)$ are added to $\state$, and those
from $\delete(\detAction)$ are removed from $\state$.

We are concerned with problems that feature partial
observability. These have a perceptual model given in terms of a
finite set of stochastic {\em senses} $\stochSenses$, deterministic
sensing outcomes $\detSenses$, and perceptual propositions
$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
a set of percepts $\observ \subseteq \percepts$, and we denote
\observations\ the set of reachable observations. The underlying state
of the process cannot be observed directly, rather, senses
$\stochSense \in \stochSenses$ effect an observation $\observ \in
\observations$ that informs what should be believed about the state a
process is in. In detail, if $\action$ is applied effecting a
transition to a successor state $\state'$, then an observation occurs
according to the active senses $\stochSenses(\stochAction, \state')
\subseteq \stochSenses$. A sense $\stochSense$ is active, written
$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
\subseteq \propositions$ is satisfied by the state $\state'$; In the
usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from chosen elements of active senses.

@@

An optimal solution to a finite-horizon POMDP is a contingent plan,
and can be expressed as a mapping from observation histories to
actions. Although suboptimal in general, useful plans can also take a
classical sequential format. This is the case in {\em conformant}
planning, where the objective is to find a sequence of actions that
achieves a goal ---i.e., reaches a state that satisfies some given
Boolean condition--- with probability $1$.  More generally, a useful
formalism for representing solutions to POMDPs is the finite-state
controller (FSC). This is a three-tuple $\langle \nodes, \selection,
\transition \rangle$ where: $\node \in \nodes$ is a set of controller
states, $\selection_\node(\action) = P(\action | \node)$ gives the
probability that the controller prescribes \action\ when in state
\node, and $\transition_\node(\action, \observation, \node') =
P(\node'|\node, \action, \observation)$ gives the probability the
controller transitions to state $\node'$, supposing we execute
\action\ in state \node\ receiving observation \observation. When
acting for finite $N$ steps from controller state $n$, the value of a
controller corresponds to the expected cumulative reward:

\begin{equation}\label{eq:expectedvalue}
V_{\pp{FSC}}(\bstate, n) = \Expect \bigg{[} 
\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{FSC}_t) \mid \pp{FSC}, n_0 = n, \bstate_0 = \bstate \bigg{]}
\end{equation}

\noindent Where $b_t$ is the belief-state at step $t$, $\pp{FSC}_t$ is
the action prescribed at step $t$, and

@@

\noindent Finally, it is
useful to note that there is a corresponding deterministic FSCs for
both sequential (e.g., conformant) and contingent plan formats.

>>>>

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. We model a process
state $\state$ as the set of propositions that are true of the
state. Notationally, we have $\state \subseteq
\props$. The underlying process
dynamics are modelled in terms of a finite set of {\em probabilistic
STRIPS operators}~\cite{boutilier:1996} $\stochActions$ over
state-characterising propositions $\propositions$.
%%
Notationally, we say an action $\stochAction \in \stochActions$ is
applicable if its precondition $\poss(\stochAction)$, a set of
propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
\state$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
nature chooses a deterministic STRIPS effect $\detAct_i$, and for
all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$.

%% We model a process state $\state$ as the set of propositions that are
%% true of the state. Notationally, we have $\state \subseteq
%% \props$. State change is induced by application of actions. A
%% stochastic action $\stochAction \in \stochActions$ is applicable if
%% its precondition $\poss(\stochAction)$, a set of propositions, are
%% satisfied in the current state. We write $\stochActions(\state)$ for set of actions
%% applicable at state $\state$.  If action $\stochAction \in
%% \stochActions(\state)$ is applied at \state, nature chooses one
%% element amongst a small set of deterministic outcomes
%% $\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
%% \}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
%% nature takes outcome $\detAct_i$, and for all \stochAction\ we require
%% $\sum_{\detAction_i \in \detActions(\stochAction)}
%% \mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
%% effect on the state given in terms of two lists of propositions. The
%% add-list $\add(\detAction)$ and delete-list
%% $\delete(\detAction)$.\footnote{If a proposition is in the add-list of
%% $\detAction$, then it cannot be in the delete-list and vice versa.}
%% If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
%% ..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
%% ..,\prop_m]$ is chosen by nature when \stochAction\ is
%% applied at state $\state$, then the resultant state is $ ( \state \cup
%% \add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
%% propositions from $\add(\detAction)$ are added to $\state$, and those
%% from $\delete(\detAction)$ are removed from $\state$.

%% \citeauthor{rintanen:01}~(\citeyear{rintanen:01})

We are concerned with problems that feature partial
observability. Although we could invoke {\em extended probabilistic
STRIPS operators}~\cite{rintanen:01} to model actions and observations
propositionally, we find it convenient for our presentation to clearly
separate sensing and action. Therefore, we suppose a POMDP has a
perceptual model given in terms of a finite set of stochastic {\em
senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
and perceptual propositions $\percepts$, called {\em percepts}. In
detail, we take an observation $\observ$ to be a set of percepts
$\observ \subseteq \percepts$, and denote \observations\ the set of
reachable observations. The underlying state of the process cannot be
observed directly, rather, senses $\stochSense \in \stochSenses$
effect an observation $\observ \in
\observations$ that informs what should be believed about the state a
process is in. In detail, if $\action$ is applied effecting a
transition to a successor state $\state'$, then an observation occurs
according to the active senses $\stochSenses(\stochAction, \state')
\subseteq \stochSenses$. A sense $\stochSense$ is active, written
$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
\subseteq \propositions$ is satisfied by the state $\state'$; In the
usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from chosen elements of active
senses.

@@

An optimal solution to a finite-horizon POMDP is a contingent plan,
and can be expressed as a mapping from observation histories to
actions. Although suboptimal in general, useful plans can also take a
classical sequential format. This is the case in {\em conformant}
planning, where the objective is to find a sequence of actions that
achieves a goal ---i.e., reaches a state that satisfies a given
Boolean condition--- with probability $1$.  Generally, whatever the
structure of a plan, its value corresponds to the expected cumulative
reward:

%% More generally, a useful
%% formalism for representing solutions to POMDPs is the finite-state
%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
%% \transition \rangle$ where: $\node \in \nodes$ is a set of controller
%% states, $\selection_\node(\action) = P(\action | \node)$ gives the
%% probability that the controller prescribes \action\ when in state
%% \node, and $\transition_\node(\action, \observation, \node') =
%% P(\node'|\node, \action, \observation)$ gives the probability the
%% controller transitions to state $\node'$, supposing we execute
%% \action\ in state \node\ receiving observation \observation. When
%% acting for finite $N$ steps from controller state $n$, the value of a
%% controller corresponds to the expected cumulative reward:

\begin{equation}\label{eq:expectedvalue}
V_{\pp{PLAN}}(\bstate, n) = \Expect \bigg{[} 
\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0 = \bstate \bigg{]}
\end{equation}

\noindent Where $b_t$ is the belief-state at step $t$, $\pp{PLAN}_t$ is
the action prescribed at step $t$, and

@@

%% \noindent Finally, it is
%% useful to note that there is a corresponding deterministic FSCs for
%% both sequential (e.g., conformant) and contingent plan formats.
